---
title: "Tidy Tuesday: Quantifying Art history data"
date: 1/17/2023
author:
  - name: Deepsha Menghani 
    url: https://github.com/deepshamenghani
title-block-banner: true
format:
  html:
    theme: flatly
    code-fold: false
    code-tools: true
    toc: false
    number-sections: true
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  eval = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.height = 7
)
```

In this post, I will analyse the [#TidyTuesday](https://github.com/rfordatascience/tidytuesday/tree/master/) Art history dataset.

[Arthistory data package](https://saralemus7.github.io/arthistory/) contains both artist-level and work-level data that was collected along with variables regarding the artists\' demographics and numeric metrics for describing how much space they or their work took up in each edition of each textbook - Janson’s History of Art and Gardner’s Art Through the Ages, two of the most popular art history textbooks used in the American education system.

::: callout-tip
You can find my github code repository [here](https://github.com/deepshamenghani/tidytuesday/tree/master/2023/Week3_ArtHistory).
:::

## Load libraries

```{r install packages, message=FALSE, warning=FALSE, include=TRUE, paged.print=FALSE, eval=TRUE, echo=TRUE}

# For loading Tidy Tuesday data
library(tidytuesdayR)

# EDA
library(tidyverse)
library(DT)

# Plotting
library(camcorder)
library(showtext)
library(ggthemes)
library(cowplot)
library(magick)
library(ggimage)

```

## Load data

```{r}

artists <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-01-17/artists.csv')

```
## EDA

```{r}
artists %>% 
  datatable()
```

## Editing images using Magick
```{r}
Gardnerimage <- image_read("GardnersATTP2.png")
Gardnerimage_transparent <- image_colorize(Gardnerimage, opacity = 60, color = "white")
Gardnerimage_transparent2 <- image_colorize(Gardnerimage, opacity = 20, color = "white")
```

## Getting font 
```{r}

font_add(family = "MedievalSharp-Regular.ttf", regular= "MedievalSharp-Regular.ttf")
showtext_auto()

```

## Updating data

```{r}
artist_spaces <- artists %>%
  filter(!is.na(artist_race_nwi)) %>% 
  filter(!(artist_gender %>% str_detect("N/A"))) %>% 
  mutate(artist_gender_race = str_c(artist_gender, " (", artist_race_nwi,")")) %>% 
  # filter((book %in% c("Gardner"))) %>%
  group_by(book, year, artist_gender_race) %>%
  summarise(maxspace = max(space_ratio_per_page_total, na.rm=TRUE),
            avgspace = mean(space_ratio_per_page_total, na.rm=TRUE),
            minspace = min(space_ratio_per_page_total, na.rm=TRUE)) %>%
  ungroup() %>% 
  mutate(artist_gender_race = as_factor(artist_gender_race)) %>% 
  mutate(label = if_else(year == max(year), as.character(artist_gender_race), NA_character_)) %>% 
  mutate(label_year = if_else(year %in% c(1936, 1959, 1986, 2016), as.character(year), NA_character_)) %>% 
  filter((book %in% c("Gardner"))) 
```

## Creating ggplot

```{r}
artist_space_plot <-  artist_spaces %>%
  ggplot(aes(year, maxspace ,color = artist_gender_race)) +
  geom_line(size = 2) +
  geom_text(aes(label = label),
            na.rm = TRUE,
            size = 15,
            hjust = -0.15,
            vjust=0.5, family = "MedievalSharp-Regular.ttf") +
  theme_minimal() +
  theme(
    panel.grid = element_blank()
  ) +
  scale_color_manual(values = c("#8c684e", "gray18", "#337b70", "yellow4")) +
  # scale_color_manual(values = c("darkgreen", "gray18", "brown", "yellow4")) +
  # scale_color_manual(values=c("#8c684e", "#337b70", "#0b0a5f", "#ce532a")) +
  scale_y_continuous(limits = c(0, 3.5), breaks = seq(0, 3.5, by = 0.5)) +
  scale_x_continuous(limits = c(1926, 2045), breaks = seq(1926, 2045, by = 30))  +
  labs(
    title = "Maximum page space by artist gender and race",
    x= "",
    y = "",
    color = "",
    subtitle = "Gardner’s Art Through the Ages"
  ) +
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(hjust = 2, size = 50),
    text = element_text(size = 2, family = "MedievalSharp-Regular.ttf", color="brown4"),
    title = element_text(family = "MedievalSharp-Regular.ttf", color="brown4",size = 50, hjust= 1, vjust=10),
    legend.position = "none"
  )  +
  geom_text(aes(label = label_year, y=0.01), 
            na.rm = TRUE,
            size = 18,
            # hjust = -0.15,
            vjust=0.7, family = "MedievalSharp-Regular.ttf", color="gray20")


```

## Superimposing with book cover

```{r}
ggdraw()  +
  draw_image(Gardnerimage_transparent, x = 0.04, height=1.05) +
  draw_plot(artist_space_plot) +
  draw_image(Gardnerimage_transparent2, x = 0.97, y = 1.04, hjust = 1, vjust = 1, width = 0.25, height = 0.4) 
```
## Saving image with proper resolution

```{r}
ggsave("Arthistory_Gardner.png",device = "png", # device to use to save images
       width = 9, # width of saved image
       height = 12, # height of saved image
       units = "in", # units for width and height
       dpi = 300)

```





